Traffic aware Resource Allocation schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} systems

### 2014-10-01 Report/conclusions.tex
double_ready

In this work, we presented the \acl{QM} objective to minimize the number of backlogged packets by designing the transmit precoders across the space-frequency dimension in a multi-cell \ac{MU} \ac{MIMO} system. We proposed the \ac{JSFRA} scheme, which adopts the \ac{SCA} technique to model the nonconvex constraint as a convex constraint in an iterative manner to design the precoders for the \acl{QM} objective. We also proposed an alternative approach using the \ac{MSE} relaxation for the same problem when the receive beamformers are based on the \ac{MMSE} receivers. The exponent used in the proposed formulation can be modified to achieve either greedy allocation or equally fair allocation. We also proposed the distributed solutions for the proposed centralized \ac{JSFRA} problem using the \ac{ADMM} method, which has better convergence behavior compared to the primal decomposition. In addition, we also proposed an iterative algorithm based on the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme to design the precoders in a distributed manner using the closed form solutions.

>>>>

In this paper, we addressed the allocation of space-frequency resources to the users in a multi-cell \ac{MIMO} \ac{OFDM} system. The resource allocation is considered as a joint space-frequency precoder design problem since the allocation of a resource to a user can be achieved by a non-zero precoding vector. We proposed the \ac{JSFRA} scheme by adopting the \ac{SCA} technique to model the nonconvex constraint as a sequence of convex subsets to design the precoders for minimizing the number of queued packets. Additionally, an alternative approach using \ac{MSE} relaxation is also proposed with the same objective by fixing the receivers based on \ac{MSE} minimization. We also proposed the distributed solutions for the centralized \ac{JSFRA} problem using \ac{ADMM} and primal decomposition methods. Finally, we proposed an iterative algorithm to determine the precoders in a decentralized manner based on the \ac{KKT} conditions for the \ac{MSE} reformulated \ac{JSFRA} scheme. Numerical results shows that the proposed algorithms perform better than the existing approaches.

### 2014-10-01 Report/decomposition_solutions.tex
double_ready

To begin with, let \me{\bar{\mc{B}}_b} denote the set \me{\mc{B} \backslash \{b\}} and \me{\bar{\mc{U}}_b} represents the set \me{\mc{U} \backslash \mc{U}_b}. In order to study the decomposition based solutions, we consider the solution proposed in the \eqref{eqn-9}, which is based on the Taylor approximation for the nonconvex constraint. The following discussions are equally valid for the \ac{MSE} based solution outlined in \eqref{eqn-mse-2} as well. Since the objective of \eqref{eqn-9} can be decoupled across each \ac{BS}, the centralized problem can be equivalently written as
\begin{IEEEeqnarray}{CCl} \label{eqn-decent-1}
\underset{\substack{\gamma_{l,k,n},\mvec{M}{k,n},\mvec{W}{k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \sum_{b \in \mc{B}} \| \tilde{\mbf{v}}_b \|_q \IEEEyessubnumber \label{eqn-decent-1a} \\
\text{subject to}&\quad& \eqref{eqn-9.1c} - \eqref{eqn-9.1e} \IEEEyessubnumber,
\end{IEEEeqnarray}
where \me{\tilde{\mbf{v}}_b} denote the vector of of weighted queue deviation corresponding to the users \me{k \in \mc{U}_b}.

>>>>

To begin with, let \me{\bar{\mc{B}}_b} denote the set \me{\mc{B} \backslash \{b\}} and \me{\bar{\mc{U}}_b} represents the set \me{\mc{U} \backslash \mc{U}_b}. In order to study the decomposition based solutions, we consider the solution proposed in the \eqref{eqn-9}, which is based on the Taylor approximation for the nonconvex constraint. The following discussions are equally valid for the \ac{MSE} based solution outlined in \eqref{eqn-mse-2} as well. Since the objective of \eqref{eqn-9} can be decoupled across each \ac{BS}, the centralized problem can be equivalently written as
{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl} \label{eqn-decent-1}
\underset{\substack{\gamma_{l,k,n},\mvec{M}{k,n},\mvec{W}{k,n}, \beta_{l,k,n}}}{\text{minimize}} &\quad & \sum_{b \in \mc{B}} \| \tilde{\mbf{v}}_b \|_q \IEEEyessubnumber \label{eqn-decent-1a} \\
\text{subject to}&\quad& \eqref{eqn-9.1c} - \eqref{eqn-9.1e} \IEEEyessubnumber,
\end{IEEEeqnarray}}
where \me{\tilde{\mbf{v}}_b} denote the vector of of weighted queue deviation corresponding to the users \me{k \in \mc{U}_b}.

### 2014-10-01 Report/introduction.tex
double_ready

The last mile wireless connectivity poses significant bottleneck in the overall data traffic for the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which impacts the mobile battery consumption. In order to overcome the receiver complexity, \ac{OFDM} is introduced for the wide band transmissions. To improve data rate, multiple antennas are installed at \acp{BS} and/or at user terminals to avail additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in wireless networks provides higher data rate or lower outage for the same transmission power and bandwidth.

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources subject to certain system limitations and constraints. We consider the problem of transmit design over the space-frequency resources provided by the \acs{MIMO}-\ac{OFDM} framework in the downlink broadcast transmission to minimize the number of queued packets. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem of interest can be viewed as a resource allocation to minimize the total number of backlogged packets of all \acp{BS}.

In general, the resource allocation problems are solved by assigning a binary variable to each user indicating the presence or the absence on a particular resource. In contrast to that, we use the transmit beamformers, which are the complex vectors, as a decision variable in determining the presence or the absence of a user on a particular resource. The purpose of using the transmit beamformers for the scheduling is two fold. Firstly, it determines the transmission rate on a certain resource and secondly, by making the transmit beamformer to be a zero vector, the corresponding user will not be scheduled on a certain resource.

In order to reduce the complexity involved in the precoder design problem, linear precoding are assumed at the \acp{BS} and detection is used at users. The queue minimizing network optimization objective is used to design the beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in the queues prior to the transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are closely related to the \ac{WSRM} problem with additional rate constraints determined by the number of backlogged packets for each user in the system. The topics on \ac{MIMO} \ac{IBC} precoder design have been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the \ac{MIMO} \ac{IBC} precoder design problems, the \ac{SCA} method has become a powerful tool to deal with these problems. For example, in \cite{sin_algorithm}, the nonconvex part of the objective is linearized around an operating point in order to solve the \ac{WSRM} problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem via arithmetic-geometric inequality is proposed in \cite{tran2012fast}.

@@

Earlier studies on the queue minimization problem was summarized in the survey paper \cite{berry2004cross}. In particular, the problem of power allocation to minimize the number of backlogged packets was considered in \cite{qps_cioffi} using geometric programming. Since the problem addressed in \cite{qps_cioffi} assumed single antenna transmitters and receivers, the queue minimizing problem reduces to the optimal power allocation problem. In the context of wireless networks, the backpressure algorithm mentioned above was extended in \cite{weeraddana2011resource} by formulating the corresponding user queues as the weights in the \ac{WSRM} problem.

In this paper, we consider the problem of precoder design across the space-frequency resources to minimize the total number of queued packets of all \acp{BS}. For this highly nonconvex problem, we first propose two centralized methods. In the first method, we relax the nonconvex constraint by the first order Taylor approximation around an operating point, which is updated in an iterative manner until convergence or to a certain accuracy. In the second method, we reformulate the \ac{JSFRA} problem using the \ac{MSE} equivalence with the rate expression to solve for the optimal precoders. For distributed implementation, we further proposed decentralized approaches based on primal and \ac{ADMM} scheme to identify the precoders independently across the \acp{BS} by exchanging limited information via back-haul. We also proposed an iterative algorithm by solving the \ac{KKT} equations, which can be implemented easily in a distributed manner.

The remainder of this paper is organized as follows. In Section \ref{sec-2-3.2}, we introduce the system model and the problem formulation for the queue minimizing precoder design. Existing and the proposed precoder designs for the \ac{JSFRA} problem are presented in Section \ref{sec-3}. The distributed solutions are provided in Section \ref{sec-4} followed by the simulation results in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.

>>>>

%The last mile wireless connectivity poses significant bottleneck in the overall data traffic for the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which impacts the mobile battery consumption. In order to overcome the receiver complexity, \ac{OFDM} is introduced for the wide band transmissions. To improve data rate, multiple antennas are installed at \acp{BS} and/or at user terminals to avail additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in wireless networks provides higher data rate or lower outage for the same transmission power and bandwidth.

In a network with multiple \acp{BS} serving \ac{MU}, the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources subject to certain system limitations and constraints. We consider the problem of transmit design over the space-frequency resources provided by the \acs{MIMO} \ac{OFDM} framework in the downlink \ac{IBC} to minimize the number of queued packets. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, it can be viewed as a resource allocation problem.

In general, the resource allocation problems are solved by assigning a binary variable to each user indicating the presence or the absence on a particular resource \cite{admission_control}. In contrast to that, we use the transmit beamformers, which are the complex vectors, as a decision variable in determining the presence or the absence of a user on a particular resource. The purpose of using the transmit beamformers for the scheduling is two fold. Firstly, it determines the transmission rate on a certain resource and secondly, by making the transmit beamformer to be a zero vector, the corresponding user will not be scheduled on a certain resource.

%The queue minimizing network optimization objective is used to design the beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in the queues prior to the transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are closely related to the \ac{WSRM} problem with additional rate constraints determined by the number of backlogged packets for each user in the system. The topics on \ac{MIMO} \ac{IBC} precoder design have been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the \ac{MIMO} \ac{IBC} precoder design problems, the \ac{SCA} method has become a powerful tool to deal with these problems \cite{bertsekas1999nonlinear}. For example, in \cite{sin_algorithm}, the nonconvex part of the objective is linearized around an operating point in order to solve the \ac{WSRM} problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem by using arithmetic-geometric inequality is proposed in \cite{tran2012fast}.

@@

Earlier studies on the queue minimization problem was summarized in the survey paper \cite{berry2004cross}. In particular, the problem of power allocation to minimize the number of backlogged packets was considered in \cite{qps_cioffi} using geometric programming. Since the problem addressed in \cite{qps_cioffi} assumed single antenna transmitters and receivers, the queue minimizing problem reduces to the optimal power allocation problem. In the context of wireless networks, the backpressure algorithm mentioned above was extended in \cite{weeraddana2011resource} by formulating the corresponding user queues as the weights in the \ac{WSRM} problem. 
Recently, the precoder design for the video transmission over \ac{MIMO} system is considered in \cite{video_queues}. In this design, the \ac{MU}-\ac{MIMO} precoders are designed by the \ac{MSE} reformulation as in \cite{christensen2008weighted} with the higher layer performance objective such as playback interruptions and buffer overflow probabilities.

In this paper, we consider the problem of precoder design across the space-frequency resources to minimize the total number of queued packets waiting in all \acp{BS}. For this highly nonconvex problem, we first propose two centralized methods. In the first method, we relax the nonconvex constraint by the first order Taylor approximation around an operating point, which is updated in an iterative manner until convergence or to a certain accuracy. In the second method, we reformulate the \ac{JSFRA} problem using the \ac{MSE} equivalence with the rate expression to solve for the optimal precoders. For distributed implementation, we further proposed decentralized approaches based on primal and \ac{ADMM} scheme to identify the precoders independently across the \acp{BS} by exchanging limited information via back-haul. We also proposed an iterative algorithm by solving the \ac{KKT} equations, which can be implemented easily in a distributed manner.

The organization of this paper is as follows. In Section \ref{sec-2-3.2}, we introduce the system model and the problem formulation for the queue minimizing precoder design. Existing and the proposed precoder designs for the \ac{JSFRA} problem are presented in Section \ref{sec-3}. The distributed solutions are provided in Section \ref{sec-4} followed by the simulation results in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.

### 2014-10-01 Report/joint_wsrm_formulation.tex
temporary_version
double_ready

The queue minimizing algorithms are discussed extensively in the networking literature to provide congestion-free routing between any two nodes in the network.\footnote{routers or user terminals} One such algorithm  is the \emph{backpressure algorithm}, discussed in detail in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes. Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

@@

&=&\frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\

@@

}{
	\underset{\mbf{t}}{\text{minimize}} && \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
	&\leq& \underbrace{\mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set {\sum_{k \in \mc{U}} \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)}}_{\le B} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) \nonumber \\
	&& - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)} \IEEEyessubnumber \label{backpreassure-approx-1b}
}
where the second term in \eqref{backpreassure-approx-1b} follows from the Poisson arrival process. Assuming the second order moments of the transmissions and the arrival rates are bounded, it can be replaced by a upper bound \me{B} as in \eqref{backpreassure-approx-1b}, in order to eliminate from the optimization problem \cite{neely2010stochastic}. The expectation is performed across all possible arrivals and the transmission rates for the given number of queued packets \me{\mbf{Q}(i)} and the channel state information at the \me{\ith{i}} slot.

>>>>

The queue minimizing algorithms are discussed extensively in the networking literature to provide congestion-free routing between any two nodes in the network\footnote{routers or user terminals}. One such algorithm  is the \emph{backpressure algorithm}, discussed in detail in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes. Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

@@

&=&\frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ] \IEEEyessubnumber \label{eqn-3.1.0} \\

@@

\end{IEEEeqnarray}}{
{\allowdisplaybreaks
\begin{IEEEeqnarray}{CL} \label{backpressure-approx}
	& \underset{\mbf{t}}{\text{minimize}} \quad \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
	\leq & \underbrace{\mathbb{E}_{\mbfa{\lambda},\mbf{t}} \Big \lbrace \sum_{k \in \mc{U}} \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i) \Big \rbrace }_{\le B} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) \nonumber \\
	& \qquad \qquad {} - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\Big \lbrace \sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i) \Big \rbrace \IEEEyessubnumber \label{backpreassure-approx-1b}
\end{IEEEeqnarray}}}
%\begin{figure*}
%\begin{IEEEeqnarray}{rCl} \label{backpressure-approx}
%	\underset{\mbf{t}}{\text{minimize}} &\quad& \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \vspace{-0.25in}\\
%	& \leq & \sum_{k \in \mc{U}} Q_k(i) A_k(i) - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)} + \overbrace{\mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set {\sum_{k \in \mc{U}} \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)}}^{\le B}\IEEEyessubnumber \label{backpreassure-approx-1b}
%\end{IEEEeqnarray}	
%\vspace{-0.25in}
%\end{figure*}
where the first term in \eqref{backpreassure-approx-1b} follows from the Poisson arrival process.

Assuming the second order moments of the transmissions and the arrival rates are bounded, it can be replaced by a upper bound \me{B} as in \eqref{backpreassure-approx-1b}, in order to eliminate from the optimization problem \cite{neely2010stochastic}. The expectation is performed across all possible arrivals and the transmission rates for the given number of queued packets \me{\mbf{Q}(i)} and the channel state information at the \me{\ith{i}} slot.

### 2014-10-01 Report/mse_formulation.tex
double_ready

The \acl{QM} \ac{JSFRA} problem using alternative \ac{MSE} formulation given by \eqref{eqn-mse-1} is non-convex even for the fixed \me{\mvec{w}{l,k,n}} due to the constraint \eqref{eqn-mse-1.2}. In order to solve this efficiently, we use the \ac{SCA} method as discussed earlier in Section \ref{sec-3.2} by using the linear under estimator for the convex function on the r.h.s of \eqref{eqn-mse-1.2}. The first order approximation around a fixed \ac{MSE} point \me{\tilde{\epsilon}_{l,k,n}} for \eqref{eqn-mse-1.2} is given by
\begin{IEEEeqnarray}{rCl}
- \log_2(\tilde{\epsilon}_{l,k,n}) - \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\log(2) \, \tilde{\epsilon}_{l,k,n}} &\geq& t_{l,k,n}
\label{mse-lin}
\end{IEEEeqnarray}

>>>>

The alternative \ac{MSE} formulation given by \eqref{eqn-mse-1} is non-convex even for the fixed \me{\mvec{w}{l,k,n}} due to the constraint \eqref{eqn-mse-1.2}. We adopt the \ac{SCA} method as in Section \ref{sec-3.2} to relax the constraint by a sequence of convex subsets using first order approximations around \me{\tilde{\epsilon}_{l,k,n}} as
\begin{IEEEeqnarray}{rCl}
- \log_2(\tilde{\epsilon}_{l,k,n}) - \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\log(2) \, \tilde{\epsilon}_{l,k,n}} &\geq& t_{l,k,n}
\label{mse-lin}
\end{IEEEeqnarray}

### 2014-10-01 Report/system_model.tex
double_ready

In this paper we adopt linear beamforming technique at \acp{BS}. Specifically, the data symbols \me{d_{l,k,n}} for  user \me{k} on the \me{l^\mathrm{th}} spatial stream over the sub-channel \me{n} is multiplied with the beamformer \me{\mvec{m}{l,k,n} \in \mathbb{C}^{N_T \times 1}} before being transmitted. In order to detect  multiple spatial streams at the receiver, a receive beamforming vector \me{\mvec{w}{l,k,n}} is employed at each user. Consequently, the received data symbol corresponding to the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n} at user $k$ is given by
\iftoggle{single_column}{
\begin{equation}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n}
\end{equation}
}{
\begin{multline}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n} \\ 
+ \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n}
\end{multline}
}
where \me{\mvec{H}{b,k,n} \in \mathbb{C}^{N_R \times N_T}} is the channel between \ac{BS} \me{b} and  user \me{k} on sub-channel \me{n}, and \me{\mvec{n}{k,n}} \me{\sim \mc{CN}(0,N_0)} is the additive noise vector for the user \me{k} on the \me{n^\mathrm{th}} sub-channel and \me{l^\mathrm{th}} spatial stream. In \eqref{eqn-1}, \me{L = rank(\mvec{H}{b,k,n}) = min(N_T,N_R)} is the maximum number of spatial streams.\footnote{\me{L} streams are initialized but after solving the problem, only \me{L_{k,n} \leq L} non-zero data streams are transmitted} Assuming independent detection of data streams, we can write the signal-to-interference-plus-noise ratio (SINR) as
\begin{equation}\label{eq:SINR}
\gamma_{l,k,n} = \dfrac{\left |\mvec{w}{l,k,n}^\herm \, \mvec{H}{b_k,k,n} \, \mvec{m}{l,k,n} \right |^2}{N_0\|\mvec{w}{l,k,n}\|^2 + \sum_{(j,i) \neq (l,k)} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2}
\end{equation}
Let \me{Q_k} be the number of backlogged packets which are destined for the user \me{k} at a given scheduling instant. The queue dynamics of the user \me{k} are modeled using the Poisson arrival process with the average packet arrivals of \me{A_k = \mathbf{E}_i\{\lambda_k\}} packets/bits, where \me{\lambda_k(i) \sim \mathrm{Pois}(A_k)} represents the instantaneous number of packets or bits arriving for the user \me{k} at the \me{{i}^\mathrm{th}} instant. The total number of queued packets at the \me{{(i+1)}^\mathrm{th}} instant for the user \me{k}, represented by \me{Q_k(i+1)}, is given by
\begin{equation}
Q_k(i+1) = \Big [ Q_k(i) - t_k(i) \Big ]^+ + \lambda_k(i)
\label{eqn-2a}
\end{equation}
where \me{[x]^+ \equiv \max{ \lbrace x,0  \rbrace}} and \me{t_k} denotes the transmission in bits for user \me{k}. For a \ac{MIMO}-\ac{OFDM} system,
\begin{equation}
t_k(i) = \sum_{n = 1}^N \, \sum_{l = 1}^L \, t_{l,k,n}(i)
\end{equation}
where \me{t_{l,k,n}} denotes the transmitted bits over \me{\ith{l}} spatial stream on the \me{\ith{n}} sub-channel. The maximum rate achieved over the \me{(l,n)} space-frequency resource is given by \me{t_{l,k,n} \leq \log_2(1 + \gamma_{l,k,n})} for the \ac{SINR} of \me{\gamma_{l,k,n}}.\footnote{This can be achieved by Gaussian signaling} Note that the units of \me{t_k} and \me{Q_k} are in bits defined per channel use.

>>>>

In this paper we adopt linear beamforming technique at \acp{BS}. Specifically, the data symbols \me{d_{l,k,n}} for  user \me{k} on the \me{l^\mathrm{th}} spatial stream over the sub-channel \me{n} is multiplied with the beamformer \me{\mvec{m}{l,k,n} \in \mathbb{C}^{N_T \times 1}} before being transmitted. In order to detect  multiple spatial streams at the receiver, a receive beamforming vector \me{\mvec{w}{l,k,n}} is employed at each user. Consequently, the received data symbol corresponding to the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n} at user $k$ is given by
\iftoggle{single_column}{
\begin{equation}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n}
\end{equation}
}{
\begin{multline}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n} \\ 
+ \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n}
\end{multline}
}
where \me{\mvec{H}{b,k,n} \in \mathbb{C}^{N_R \times N_T}} is the channel between \ac{BS} \me{b} and  user \me{k} on sub-channel \me{n}, and \me{\mvec{n}{k,n}} \me{\sim \mc{CN}(0,N_0)} is the additive noise vector for the user \me{k} on the \me{n^\mathrm{th}} sub-channel and \me{l^\mathrm{th}} spatial stream. In \eqref{eqn-1}, \me{L = rank(\mvec{H}{b,k,n}) = min(N_T,N_R)} is the maximum number of spatial streams\footnote{\me{L} streams are initialized but after solving the problem, only \me{L_{k,n} \leq L} non-zero data streams are transmitted}. Assuming independent detection of data streams, we can write the signal-to-interference-plus-noise ratio (SINR) as
\begin{equation}\label{eq:SINR}
\gamma_{l,k,n} = \dfrac{\left |\mvec{w}{l,k,n}^\herm \, \mvec{H}{b_k,k,n} \, \mvec{m}{l,k,n} \right |^2}{N_0\|\mvec{w}{l,k,n}\|^2 + \sum_{(j,i) \neq (l,k)} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2}
\end{equation}
Let \me{Q_k} be the number of backlogged packets which are destined for the user \me{k} at a given scheduling instant. The queue dynamics of the user \me{k} are modeled using the Poisson arrival process with the average packet arrivals of \me{A_k = \mathbf{E}_i\{\lambda_k\}} packets/bits, where \me{\lambda_k(i) \sim \mathrm{Pois}(A_k)} represents the instantaneous number of packets or bits arriving for the user \me{k} at the \me{{i}^\mathrm{th}} instant. The total number of queued packets at the \me{{(i+1)}^\mathrm{th}} instant for the user \me{k}, represented by \me{Q_k(i+1)}, is given by
\begin{equation}
Q_k(i+1) = \Big [ Q_k(i) - t_k(i) \Big ]^+ + \lambda_k(i)
\label{eqn-2a}
\end{equation}
where \me{[x]^+ \equiv \max{ \lbrace x,0  \rbrace}} and \me{t_k} denotes the transmission in bits for user \me{k}. For a \ac{MIMO}-\ac{OFDM} system,
\begin{equation}
t_k(i) = \sum_{n = 1}^N \, \sum_{l = 1}^L \, t_{l,k,n}(i)
\end{equation}
where \me{t_{l,k,n}} denotes the transmitted bits over \me{\ith{l}} spatial stream on the \me{\ith{n}} sub-channel. The maximum rate achieved over the \me{(l,n)} space-frequency resource is given by \me{t_{l,k,n} \leq \log_2(1 + \gamma_{l,k,n})} for the \ac{SINR} of \me{\gamma_{l,k,n}}\footnote{This can be achieved by Gaussian signaling}. Note that the units of \me{t_k} and \me{Q_k} are in bits defined per channel use.
